---
title: "ggplot2: 11 Colour Scales and Legends"
output: html_notebook
---

This notebook is based on exercises completed from [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html). 

```{r setup, warning=FALSE,message=FALSE}
library(tidyverse)

```

## 11.3.4 Exercise

Recreate the following plot: (see website)

```{r}

  ggplot(mpg, aes(displ, hwy, color = drv)) +
  geom_point(size = 3) +
  scale_color_discrete(labels = c("4wd", "fwd", "rwd"))


```


## 11.7.5 Exercises

1. How do you make legends appear to the left of the plot?

```{r}

# given
toy <- data.frame(
  const = 1, 
  up = 1:4,
  txt = letters[1:4], 
  big = (1:4)*1000,
  log = c(2, 5, 10, 2000)
)

base <- ggplot(toy, aes(up, up)) + 
  geom_point(aes(colour = txt), size = 3) + 
  xlab(NULL) + 
  ylab(NULL)

base + theme(legend.position = "left")

```


2. What’s gone wrong with this plot? How could you fix it?

There are two legends. I would combine them into one for shape and color

```{r}

# given
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = drv, shape = drv)) + 
  scale_colour_discrete("Drive train")

# fixed
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = drv, shape = drv)) + 
  labs(shape = "Drive train", colour = "Drive train")

# # alternative
# ggplot(mpg, aes(displ, hwy)) + 
#   geom_point(aes(colour = drv, shape = drv)) + 
#   scale_colour_discrete("Drive train") +
#   scale_shape_discrete("Drive train")
# # not as good since returns NAs


```


3. Can you recreate the code for this plot?

Still need to drop outer levels using drop limits and override.aes??

```{r}

ggplot(mpg, aes(displ, hwy, colour = class)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, formula = y~x) +

  # to put legend in single row
  guides(colour = guide_legend(nrow = 1,
                               override.aes = list(linetype = c(0,1,1,1,1,1,0)))) +
  # to put legend at bottom
  theme(legend.position = "bottom",
        # to remove legend title
        legend.title = element_blank())

# likely need to drop limits instead to remove 2seater and SUV from legend
# how can we do this??

```

