---
title: "ggplot2: 14 Build a plot layer by layer"
output: html_notebook
---

This notebook is based on exercises completed from [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html). 

```{r setup, warning=FALSE,message=FALSE}
library(tidyverse)

```

## 14.3.1 Data Exercises

1. The first two arguments to ggplot are data and mapping. The first two arguments to all layer functions are mapping and data. Why does the order of the arguments differ? (Hint: think about what you set most commonly.)

They are in order of most commonly used. In ggplot2, the data comes first and in geoms, usually the data is default to the ggplot call so the mappings are used most often.


2. The following code uses dplyr to generate some summary statistics about each class of car. Use the data to recreate this plot:

```{r}

class <- mpg %>% 
  group_by(class) %>% 
  summarise(n = n(), hwy = mean(hwy),
            .groups = "drop")

# create annotation vector
classMeans <- str_c("n=", class$n, sep = "") 

ggplot() +
  geom_jitter(aes(class, hwy), data = mpg) +
  geom_point(aes(class, hwy), data = class, color = "red", size = 7) +
  geom_text(aes(class, min(mpg$hwy), label = classMeans), data = class)

# alternative version from later in chapter
ggplot(mpg, aes(class, hwy)) +
  geom_jitter() +
  geom_point(stat = "summary", fun = "mean", colour = "red", size = 7)

# AM solution with label as new column
class <- mpg %>%
  group_by(class) %>%
  summarise(n = n(), hwy = mean(hwy),
            .groups = "drop") %>%
  mutate(nlabel = paste("n = ",n))

ggplot(mpg, aes(class, hwy)) +
  geom_point(position = "jitter") +
  geom_point(data = class,colour = "red",size = 5) +
  geom_text(data = class,aes(label = nlabel, y = 10))

#YX solution
t<-paste("n=",class$n)
ggplot(mpg,aes(x=class,y=hwy))+
geom_jitter(width = 0.1, size=2)+ # use width to constrain the points
geom_point(aes(x=class,y=hwy),data = class,colour="red",size=5)+
geom_text(aes(label=t,y=10),data=class)

```


## 14.4.3 Aesthetics Exercises

1. Simplify the following plot specifications:

```{r}

# given
ggplot(mpg) + 
  geom_point(aes(mpg$displ, mpg$hwy))

# fixed
ggplot(mpg) + 
  geom_point(aes(displ, hwy))

# given
ggplot() + 
 geom_point(mapping = aes(y = hwy, x = cty), data = mpg) +
 geom_smooth(data = mpg, mapping = aes(cty, hwy))

# simplified
ggplot(mpg, aes(cty, hwy)) + 
 geom_point() +
 geom_smooth()

# given
ggplot(diamonds, aes(carat, price)) + 
  geom_point(aes(log(brainwt), log(bodywt)), data = msleep)

# simlipfied
ggplot() + 
  geom_point(aes(log(brainwt), log(bodywt)), msleep)



```


2. What does the following code do? Does it work? Does it make sense? Why/why not?

```{r}

ggplot(mpg) +
  geom_point(aes(class, cty)) + 
  geom_boxplot(aes(trans, hwy))

```

This code runs so it works; however, it does not make sense because the x axis has two different values depending on the geom. Geom_point uses class, whereas geom_point uses transmission.


3. What happens if you try to use a continuous variable on the x axis in one layer, and a categorical variable in another layer? What happens if you do it in the opposite order?

The is an error generates if you try to apply different types of data to the same x-axis.

```{r}

# categorical x axis with continuous x axis on second layer
# gives error
ggplot(mpg) +
  geom_point(aes(class, hwy)) +
  geom_point(aes(hwy, class), color = "red")

```


```{r}

# continuous x axis with categorical x axis on second layer
# gives error
ggplot(mpg) +
  geom_point(aes(hwy, class), color = "red") +
  geom_point(aes(class, hwy))

```


## 14.5.1 Geoms Exercises

1. Download and print out the ggplot2 cheatsheet from http://www.rstudio.com/resources/cheatsheets/ so you have a handy visual reference for all the geoms.

2. Look at the documentation for the graphical primitive geoms. Which aesthetics do they use? How can you summarise them in a compact form?

Graphical primitives use x and y as the minimum required aesthetics. They have additional optional arguments.
geom_blank()
geom_curve()
geom_path()
geom_polygon()
geom_rect()
geom_ribbon()

Compact form: aes(x,y)

3. What’s the best way to master an unfamiliar geom? List three resources to help you get started.

ggplot cheatsheet
google
?functionName for help file
RStudio help
stack overflow
ask/email a friend

4. For each of the plots below, identify the geom used to draw it.

geom_violin()
geom_point() with size as aes and alpha = 0.5?
geom_hex()
geom_jitter()
geom_area()
geom_path()

```{r}

# from YX solution
ggplot(data=mpg,aes(x=drv,y=displ))+
geom_violin(lwd=1)

ggplot(data=mpg,aes(x=hwy,y=cty))+
geom_jitter(alpha=0.2,size=mpg$displ,height = 0.1) #??

ggplot(data=mpg,aes(x=hwy,y=cty))+
geom_hex(bins=10,show.legend = F)

ggplot(data=mpg,aes(x=cyl,y=drv))+
geom_jitter(width = 0.1,height=0.2, size=3)

ggplot(data=economics,aes(x=date,y=psavert))+
geom_area()

ggplot(data=economics,aes(x=uempmed,y=psavert))+
geom_path(lwd=1)

```


5. For each of the following problems, suggest a useful geom:

-Display how a variable has changed over time.
geom_line using time on x axis
geom_step
geom_area

-Show the detailed distribution of a single variable.
geom_histogram
geom_violin
(geom_point with color aes() ->assuming you have x y coordinates)
)
-Focus attention on the overall trend in a large dataset.
geom_smooth() over geom_point
geom_boxplot or geom_violin

-Draw a map.
geom_map() or ggmap()
library(bcmaps) has high and low resolution 

-Label outlying points.
geom_point() on outlier dataset with labels or geom_label()
geom_text mapped to outliers
geom_points for outliers in bigger halo (biiger size than df)



##14.6.2 Stats Exercises

1. The code below creates a similar dataset to stat_smooth(). Use the appropriate geoms to mimic the default geom_smooth() display.

```{r}

# given
mod <- loess(hwy ~ displ, data = mpg)
smoothed <- data.frame(displ = seq(1.6, 7, length = 50))
pred <- predict(mod, newdata = smoothed, se = TRUE) 
smoothed$hwy <- pred$fit
smoothed$hwy_lwr <- pred$fit - 1.96 * pred$se.fit
smoothed$hwy_upr <- pred$fit + 1.96 * pred$se.fit

# make geom_smooth
plot1 <- ggplot(mpg, aes(displ, hwy)) +
  geom_smooth(method = "loess", formula = "y~x")
plot1

# find colors in ggplot
# this is really cool!!!
ggplot_build(plot1)$data

# make equivalent to geom_smooth
ggplot(smoothed, aes(displ, hwy)) +
  geom_ribbon(aes(ymin = hwy_lwr, ymax = hwy_upr), alpha = 0.2) +
    geom_line(color = "#3366FF", size = 1)



```

2. What stats were used to create the following plots?


3. Read the help for stat_sum() then use geom_count() to create a plot that shows the proportion of cars that have each combination of drv and trans.

## 14.7.1 Position Adjustments Exercises 

1. When might you use position_nudge()? Read the documentation.

2. Many position adjustments can only be used with a few geoms. For example, you can’t stack boxplots or errors bars. Why not? What properties must a geom possess in order to be stackable? What properties must it possess to be dodgeable?

3. Why might you use geom_jitter() instead of geom_count()? What are the advantages and disadvantages of each technique?

4. When might you use a stacked area plot? What are the advantages and disadvantages compared to a line plot?


