---
title: "ggplot2: 15 Scales and Guides"
output: html_notebook
---

This notebook is based on exercises completed from [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html). 

```{r setup, warning=FALSE,message=FALSE}
library(tidyverse)

```

## 15.1.1 Scale Specification Exercises

1.Simplify the following plot specifications to make them easier to understand.

```{r}
# given
ggplot(mpg, aes(displ)) + 
  scale_y_continuous("Highway mpg") + 
  scale_x_continuous() +
  geom_point(aes(y = hwy))

# simplified
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() +
  scale_y_continuous("Highway mpg")

# given
ggplot(mpg, aes(y = displ, x = class)) + 
  scale_y_continuous("Displacement (l)") + 
  scale_x_discrete("Car type") +
  scale_x_discrete("Type of car") + 
  scale_colour_discrete() + 
  geom_point(aes(colour = drv)) + 
  scale_colour_discrete("Drive\ntrain")

# simplify
ggplot(mpg, aes(class, displ, colour = drv)) + 
  geom_point() + 
  labs(x = "Type of car", y = "Displacement (l)") +
  scale_colour_discrete("Drive\ntrain")

```


2. What happens if you pair a discrete variable with a continuous scale? What happens if you pair a continuous variable with a discrete scale?

Error: Discrete value supplied to continuous scale

But the continuous variable appears fine with discrete scale


```{r}

# # discrete variable with continuous scale
# ggplot(mpg, aes(class)) +
#   geom_bar() +
#   scale_x_continuous()

# continuous variable to discrete scale
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() +
  scale_x_discrete()

```

## 15.7 Legend Merging and Splitting Exercises

1. What are the three most important arguments that apply to both axes and legends? What do they do? Compare and contrast their operation for axes vs. legends.

name - label or title  
breaks - ticks and grid line for key  
labels - tick or key label  

These arguments work on both axes and legends. The main difference is where they display.


## No exercises within 16 Coordinate systems - reading only

16.2.2 bullseye chart  
16.2 and 16.2.3 
Munching slows coord_map() compared to coord_quickmap
