---
title: "ggplot2: 15 Scales and Guides"
output: html_notebook
---

This notebook is based on exercises completed from [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html). 

```{r setup, warning=FALSE,message=FALSE}
library(tidyverse)

```

## 15.1.1 Scale Specification Exercises

1.Simplify the following plot specifications to make them easier to understand.

```{r}
# given
ggplot(mpg, aes(displ)) + 
  scale_y_continuous("Highway mpg") + 
  scale_x_continuous() +
  geom_point(aes(y = hwy))

ggplot(mpg, aes(y = displ, x = class)) + 
  scale_y_continuous("Displacement (l)") + 
  scale_x_discrete("Car type") +
  scale_x_discrete("Type of car") + 
  scale_colour_discrete() + 
  geom_point(aes(colour = drv)) + 
  scale_colour_discrete("Drive\ntrain")

```


2. What happens if you pair a discrete variable with a continuous scale? What happens if you pair a continuous variable with a discrete scale?

## 15.7 Legend Merging and Splitting Exercises

1. What are the three most important arguments that apply to both axes and legends? What do they do? Compare and contrast their operation for axes vs. legends.


## No exercises within 16 Coordinate systems - reading only


