---
title: "ggplot2: 18 Themes"
output: html_notebook
---

This notebook is based on exercises completed from [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html). 

```{r setup, warning=FALSE,message=FALSE}
library(tidyverse)

```

## 18.2.1 Complete Theme Exercises

1. Try out all the themes in ggthemes. Which do you like the best?

[ggtheme examples](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)

[theme_map] (https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/theme_map/)

2. What aspects of the default theme do you like? What don’t you like? What would you change?

I prefer theme_bw(). In particular, I always like to change the grey background.

```{r}

#default theme
ggplot() + 
 geom_point(mapping = aes(y = hwy, x = cty), data = mpg) 

#HDB theme
ggplot() + 
 geom_point(mapping = aes(y = hwy, x = cty), data = mpg) +
  theme_cowplot()

# EA favourite theme
ggplot() + 
 geom_point(mapping = aes(y = hwy, x = cty), data = mpg) +
  theme_bw()

# YX recommends save plot as png with background as NA
# BG = NA

```


3. Look at the plots in your favourite scientific journal. What theme do they most resemble? What are the main differences?

[Marine Ecological Progress Series](https://www.int-res.com/abstracts/meps/v666/)

[Example Article](https://www.int-res.com/articles/feature/m666p001.pdf)

On page 4,the map is simple with white background, no ticks. On page 8 the graphs are close to them_bw() with a white background and enclosed frame.

## 18.4.6 Theme Elements Exercises

1. Create the ugliest plot possible! (Contributed by Andrew D. Steen, University of Tennessee - Knoxville)

```{r}

ggplot() + 
 geom_point(mapping = aes(y = hwy, x = cty, color = class), data = mpg) +
  ggthemes::theme_excel() +
  labs(title = "My Ugly Plot") +
  theme(axis.title = element_text(family = "serif"),
        axis.text = element_text(angle = 180),
        plot.background = element_rect(fill = "yellow"),
        panel.background = element_rect(fill = "purple"),
        title = element_text(hjust = 1))
  
```


2. theme_dark() makes the inside of the plot dark, but not the outside. Change the plot background to black, and then update the text settings so you can still read the labels.

```{r}

# base plot
ggplot() + 
 geom_point(mapping = aes(y = hwy, x = cty), data = mpg) +
  theme_dark()

# change plot background to black
# change text color to read anyway
ggplot() + 
 geom_point(mapping = aes(y = hwy, x = cty), data = mpg) +
  theme_dark() +
  theme(plot.background = element_rect("black"),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white"))

```


3. Make an elegant theme that uses “linen” as the background colour and a serif font for the text.

```{r}

ggplot() + 
 geom_point(mapping = aes(y = hwy, x = cty), data = mpg) +
  labs(title = "Serif font with Linen Background") +
  theme_bw() +
  theme(plot.background = element_rect("linen"),
        text = element_text(family = "serif"))

```


4. Systematically explore the effects of hjust when you have a multiline title. Why doesn’t vjust do anything?

The vjust is constrained but the background of the plot

```{r}

p <- ggplot() + 
 geom_point(mapping = aes(y = hwy, x = cty), data = mpg) +
  labs(title = "This is a really long title \nso keep going.") 

# baseplot
p

# plot with hjust
p +
  theme(plot.title = element_text(hjust = 0.5))

p +
  theme(plot.title = element_text(hjust = 1))

# try vjust
p +
  theme(plot.title = element_text(vjust = 1))

p +
  theme(plot.title = element_text(vjust = 2))

```

