---
title: "ggplot2: 2 First Steps"
output: html_notebook
---

This notebook is based on exercises completed from [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html). 

```{r setup, warning=FALSE,message=FALSE}
library(tidyverse)
#library(ggplot2)

# set global theme for graphs
theme_set(theme_bw())

```


## 2.2.1 Fuel Economy Data

1.List five functions that you could use to get more information about the mpg data set.

```{r}
summary(mpg)
length(mpg)
str(mpg)
head(mpg)
tail(mpg)
names(mpg)
nrow(mpg)
class(mpg)
glimpse(mpg)

# opens data in RStudio, but not visible in R notebook
#View(mpg)

```


2.How can you find out what other data sets are included with ggplot2?

```{r}
data(package = "ggplot2")
# opens window in RStudio and lists the associated data
```

Data sets in package ‘ggplot2’:

diamonds                     Prices of over 50,000 round cut diamonds  
economics                    US economic time series  
economics_long               US economic time series  
faithfuld                    2d density estimate of Old Faithful data  
luv_colours                  'colors()' in Luv space  
midwest                      Midwest demographics  
mpg                          Fuel economy data from 1999 to 2008 for 38 popular models
                             of cars  
msleep                       An updated and expanded version of the mammals sleep
                             dataset  
presidential                 Terms of 11 presidents from Eisenhower to Obama  
seals                        Vector field of seal movements  
txhousing                    Housing sales in TX  


3.Apart from the US, most countries use fuel consumption (fuel consumed over fixed distance) rather than fuel economy (distance travelled with fixed amount of fuel). How could you convert cty and hwy into the European standard of l/100km?

Conversion values from [this website](https://www.asknumbers.com/mpg-to-L100km.aspx#:~:text=To%20convert%20mpg%20to%20liters,9.8%20liters%20per%20100%20km)  

```{r}

mpg %>%
  mutate(cty_metric = round(235.214/cty,0),
         hwy_metric = round(235.214/hwy,0)) %>%
  select(cty, cty_metric, hwy, hwy_metric)  

mpg %>%
    mutate(ecty = (3.78541*100)/(1.609344*cty)) %>%
    mutate(ehwy = (3.78541*100)/(1.609344*hwy)) %>%
    select(cty, ecty, hwy, ehwy) 

mpg %>%
 mutate(
 cty_euro = 235.214583/cty,
 hwy_euro = 235.214583/hwy
 ) %>%
  select(cty, cty_euro, hwy, hwy_euro)

```


4.Which manufacturer has the most models in this dataset? Which model has the most variations? Does your answer change if you remove the redundant specification of drive train (e.g. “pathfinder 4wd”, “a4 quattro”) from the model name?

Dodge has the most models in the original data set.  
The model with the most variation is caravan 2wd.  
If we remove the drive train from the model, Toyota is the manufacturer with the most models.  

```{r}
# most models by manufacturer
mpg %>%
  group_by(manufacturer) %>%
  count() %>%
  arrange(-n)

# JN version assuming year not unique to model 
mpg %>%
 count(manufacturer, sort = T)

mpg %>%
 group_by(manufacturer) %>%
 summarise(num.models = length(unique(model))) %>%
 arrange(desc(num.models))

# ST version
mpg %>%
 group_by(manufacturer) %>%
 summarise(n_distinct(model))

# CG version
mpg %>% 
  group_by(model) %>% 
  tally()

# most variation in a model defined as the most entries per one model
mpg %>%
  group_by(model) %>%
  count() %>%
  arrange(-n)

# BH version in base R
with(mpg,tapply(model,model,length))

# remove redundant specification in model name, then recount models per manufacturer
mpg %>%
  mutate(model_simple = str_replace_all(model, c("quattro" = "", "4wd" = "",
                                              "2wd" = "", "awd" = ""))) %>%
  distinct(manufacturer, model_simple, year) %>%
  group_by(manufacturer) %>%
  count() %>%
  arrange(-n)

mpg %>%
  mutate(model_simple = str_replace_all(model, c("quattro" = "", "4wd" = "",
                                              "2wd" = "", "awd" = ""))) %>%
  distinct(manufacturer, model_simple) %>%
  group_by(manufacturer) %>%
  count() %>%
  arrange(-n)

# JN version
# to get rid of drive train complication
sort(unique(mpg$model)) #looks like first 2 letters should be enough to split

 mpg %>%
   transform(model, base = substr(model, 1, 2)) %>%
   group_by(manufacturer) %>%
   summarise(num.models = length(unique(base))) %>%
   arrange(desc(num.models))
 
 list(unique(mpg$model[mpg$manufacturer=="audi"])) #the above method took audi from 3 ->2, seems to work
 
 list(unique(mpg$model[mpg$manufacturer=="toyota"])) #the above method took audi from 6 ->5, seems to work
 
 list(unique(mpg$model[mpg$manufacturer=="dodge"])) #the above method didn't affect Dodge, seems to work
 
```

### 2.3 Key Components

1. How would you describe the relationship between cty and hwy? Do you have any concerns about drawing conclusions from that plot?

As fuel economy goes up in city driving, it also goes up for highway driving. Positive, linear relationship. Concerns? Variance at tails and likely rounding giving artificial lines. Function of speed...

```{r}

ggplot(mpg, aes(cty, hwy)) +
  geom_point() +
  geom_smooth(method = "lm")

```


2. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show? Is it useful? How could you modify the data to make it more informative?

Each model only belongs to one manufacturer so not useful. And some points hidden Perhaps count models by manufacturer?

```{r}

ggplot(mpg, aes(model, manufacturer)) + 
    geom_point() +
  # added so I could read text on x-axis
    theme(axis.text.x = element_text(angle = 90))

ggplot(mpg, aes(manufacturer)) + 
    geom_bar() +
    theme(axis.text.x = element_text(angle = 90)) 

#AP version
df <- mpg %>%
 transmute("man_mod" = paste(manufacturer, model, sep = " "))

ggplot(df, aes(man_mod)) +
 geom_bar() +
 coord_flip()

```


3. Describe the data, aesthetic mappings and layers used for each of the following plots. You’ll need to guess a little because you haven’t seen all the datasets and functions yet, but use your common sense! See if you can predict what the plot will look like before running the code.

```{r}

ggplot(mpg, aes(cty, hwy)) + 
  geom_point()

ggplot(diamonds, aes(carat, price)) +
  geom_point()

ggplot(economics, aes(date, unemploy)) + 
  geom_line()

ggplot(mpg, aes(cty)) + 
  geom_histogram(bins = 30)

ggplot(mpg, aes(cty)) + 
  geom_histogram(binwidth = 1)

```

## 2.4.1 Colour, size, shape and other aesthetic attributes

1. Experiment with the colour, shape and size aesthetics. What happens when you map them to continuous values? What about categorical values? What happens when you use more than one aesthetic in a plot?

Two legends are produced if you use >1 aesthetic in a plot.

```{r}

# color continuous
mpg %>%
  ggplot(aes(displ, hwy, color = year)) +
  geom_point()

# color categorical
mpg %>%
  ggplot(aes(displ, hwy, color = class)) +
  geom_point()

# # shape continuous
# mpg %>%
#   ggplot(aes(displ, hwy, shape = year)) +
#   geom_point()
# # gives following error
# # Error: A continuous variable can not be mapped to shape

# shape categorical
mpg %>%
  ggplot(aes(displ, hwy, shape = as.factor(cyl))) +
  geom_point()

# size continuous
mpg %>%
  ggplot(aes(displ, hwy, size = cyl)) +
  geom_point()

# size categorical
mpg %>%
  ggplot(aes(displ, hwy, size = class)) +
  geom_point()

# combine two aethetics
mpg %>%
  ggplot(aes(displ, hwy, color = class, size = cty)) +
  geom_point()

```


2. What happens if you map a continuous variable to shape? Why? What happens if you map trans to shape? Why?

A continuous value cannot be mapped to shape  
Mapping trans to shape gives a warning since it has too many values to be easily visible: 	
The shape palette can deal with a maximum of 6 discrete values because more than 6 becomes difficult to discriminate; trans has 10. Consider specifying shapes manually if you must have them.

```{r}

# # shape continuous
# mpg %>%
#    ggplot(aes(displ, hwy, shape = year)) +
#    geom_point()
# # gives following error
# # Error: A continuous variable can not be mapped to shape

# trans to shape
mpg %>%
  ggplot(aes(displ, hwy, shape = trans)) +
  geom_point()


```


3. How is drive train related to fuel economy? How is drive train related to engine size and class?

Front wheel drive vehicles generally have better fuel economy than rear or 4 wheel drive.  
Engine size is generally larger for 4 wheel drive vehicles which are generally larger types of vehicles such as trucks or SUV. Engine size is smaller for front wheel drive vehicles which are generally smaller engines. Rear wheel drive have larger engines either 2 seater sports cars, SUVs, and 4 sub compacts which larger engines.

```{r}

# drive train to fuel economy
mpg %>%
  ggplot(aes(hwy, cty, color = drv)) +
  geom_point() 

# drive train to engine size and class
mpg %>%
  ggplot(aes(displ, class, color = drv)) +
  geom_point()

```


## 2.5.1 Facetting

1. What happens if you try to facet by a continuous variable like hwy? What about cyl? What’s the key difference?

Facetting is most useful with categorical variable. A continuous variable, such as hwy, has too many graphs with too few points. Using cyl gives a more informative set of graphs.


```{r}

# without facetting
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point()

# try facetting by a continuous variable
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~hwy)

# try facetting by a continuous variable
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~cyl)

```


2. Use facetting to explore the 3-way relationship between fuel economy, engine size, and number of cylinders. How does facetting by number of cylinders change your assessement of the relationship between engine size and fuel economy?

As engines get bigger the fuel efficiency drops within each number of cylinders. The number of cylinders and engine displacement are related and go up together.

```{r}

# without facetting
mpg %>%
  ggplot(aes(displ, hwy, colour = as.factor(cyl))) +
  geom_point()

# with facetting
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~cyl)

```


3. Read the documentation for facet_wrap(). What arguments can you use to control how many rows and columns appear in the output?

```{r}
#?facet_wrap

# example graph with number of columns controlled
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~cyl, nrow = 1)

# example graph with number of rows controlled
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~cyl, ncol = 1)

```


4. What does the scales argument to facet_wrap() do? When might you use it?

The scales arguement allows the axis (either x, y or both) to be different between graphs. The default is to fix all scales to the same range. This makes is harder to compare graphs, but may be necessary if the scales are very different between graphs so see the nuances.

```{r}

# example with scales argument
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~cyl, scales = "free_y")

# example CK
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~cyl, scales = "free")

```


## 2.6.6 Plot Geoms

1. What’s the problem with the plot created by ggplot(mpg, aes(cty, hwy)) + geom_point()? Which of the geoms described above is most effective at remedying the problem?

2. One challenge with ggplot(mpg, aes(class, hwy)) + geom_boxplot() is that the ordering of class is alphabetical, which is not terribly useful. How could you change the factor levels to be more informative?

  Rather than reordering the factor by hand, you can do it automatically based on the data: ggplot(mpg, aes(reorder(class, hwy), hwy)) + geom_boxplot(). What does reorder() do? Read the documentation.

3. Explore the distribution of the carat variable in the diamonds dataset. What binwidth reveals the most interesting patterns?

4. Explore the distribution of the price variable in the diamonds data. How does the distribution vary by cut?

5. You now know (at least) three ways to compare the distributions of subgroups: geom_violin(), geom_freqpoly() and the colour aesthetic, or geom_histogram() and facetting. What are the strengths and weaknesses of each approach? What other approaches could you try?

6. Read the documentation for geom_bar(). What does the weight aesthetic do?

7. Using the techniques already discussed in this chapter, come up with three ways to visualise a 2d categorical distribution. Try them out by visualising the distribution of model and manufacturer, trans and class, and cyl and trans.
  
## 2.7 Modifying the axes

No exercises

## 2.8 Output

No exercises
