---
title: "ggplot2: 3 Individual Geoms"
output: html_notebook
---

This notebook is based on exercises completed from [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html). 

```{r setup, warning=FALSE,message=FALSE}

library(tidyverse)

# set global theme for graphs
theme_set(theme_bw())

```

#3.1.1 Individual Geoms Exercises

1. What geoms would you use to draw each of the following named plots?

  1.Scatterplot - geom_point()
  2.Line chart - geom_line
  3.Histogram - geom_histogram()
  4.Bar chart - geom_bar(stat = "identity")
  5.Pie chart - geom_bar() + coord_polar() see example
  
```{r}

# from https://ggplot2.tidyverse.org/reference/coord_polar.html
pie <- ggplot(mtcars, aes(x = factor(1), fill = factor(cyl))) +
 geom_bar(width = 1)
pie + coord_polar(theta = "y")

```
  

2. What’s the difference between geom_path() and geom_polygon()? What’s the difference between geom_path() and geom_line()?

-geom_path draws a line in the order of the data whereas geom_polygon fills in the area between lines

-geom_path draws the line in the order of the data whereas geom_line draws from left to right  

3. What low-level geoms are used to draw geom_smooth()? What about geom_boxplot() and geom_violin()?

-geom_smooth uses geom_line for the best fit line and geom_area for the standard error shading
-geom_boxplot uses geom_rect or geom_tile
-geom_violin uses geom_polygon  

