---
title: "ggplot2: 5 Statistical Summaries"
output: html_notebook
---

This notebook is based on exercises completed from [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html). 

```{r setup, warning=FALSE,message=FALSE}

library(tidyverse)

# set global theme for graphs
theme_set(theme_bw())

```

# 5.4.1 Displaying Distributions Executrices

1.What binwidth tells you the most interesting story about the distribution of carat?

500 shows the market divisions of 0.25, 0.5, 0.75 1.0 etc.

```{r}

# default
diamonds %>%
  ggplot(aes(carat)) +
  geom_histogram(bins = 30) +
  labs(subtitle = "Default 30 bins")

# more bins
diamonds %>%
  ggplot(aes(carat)) +
  geom_histogram(bins = 100)  +
  labs(subtitle = "100 bins")

# less bins
diamonds %>%
  ggplot(aes(carat)) +
  geom_histogram(bins = 10)  + # 
  labs(subtitle = "10 bins")

# even more bins
diamonds %>%
  ggplot(aes(carat)) +
  geom_histogram(bins = 500)  + # or binwidth 0.01?
  labs(subtitle = "Best with 500 bins")

```


2.Draw a histogram of price. What interesting patterns do you see?

```{r}

diamonds %>%
  ggplot(aes(price)) +
  geom_histogram(bins = 1000) +
  xlim(400, 2000) +
  labs(subtitle = "Where are all the $1500 diamonds?")

diamonds %>%
  ggplot(aes(price)) +
  facet_wrap(~cut) +
  geom_histogram(bins = 1000) +
  #xlim(400, 5000) +
  labs(subtitle = "More higher quality cuts than fair")

# Christine found bimodal distribution
ggplot(diamonds, aes(price)) +
  geom_histogram(binwidth = 500)
```


3.How does the distribution of price vary with clarity?

```{r}

diamonds %>%
  ggplot(aes(price, fill = clarity, color = clarity)) +
  geom_freqpoly()

diamonds %>%
  ggplot(aes(price, fill = clarity, color = clarity)) +
  geom_histogram(bins = 500, position = "fill", na.rm = TRUE)

diamonds %>%
  ggplot(aes(price, clarity)) +
  geom_boxplot()

diamonds %>%
  ggplot(aes(price, clarity)) +
  geom_violin()

# JN used faceted version
ggplot(diamonds, aes(price)) + 
  geom_histogram(binwidth = 100) + 
  facet_wrap(~clarity)


```


4.Overlay a frequency polygon and density plot of depth. What computed variable do you need to map to y to make the two plots comparable? (You can either modify geom_freqpoly() or geom_density().)

```{r}

diamonds %>%
  ggplot(aes(depth)) +
  geom_freqpoly(binwidth = 0.1, size = 5) +
  geom_density(color = "darkred", fill = "darkred", size = 3) +
  labs(subtitle = "Original overlay")
  
diamonds %>%
  ggplot(aes(depth)) +
  geom_freqpoly(binwidth = 0.1, size = 5,
                aes(y = ..density..)) +
  geom_density(color = "darkred", fill = "darkred", alpha = 0.5) +
  labs(subtitle = "Geom freqpoly with aes(y = ..density..)")

# https://stackoverflow.com/questions/37404002/geom-density-to-match-geom-histogram-binwitdh
diamonds %>%
  ggplot(aes(depth)) +
  geom_freqpoly(binwidth = 0.1, size = 5) +
  geom_density(aes(y = 0.1 * ..count..),
               color = "darkred", fill = "darkred", alpha = 0.5) +
  labs(subtitle = "Geom denisty with aes(y =binwidth * ..count..)")
  

# JN changed the stat instead
ggplot(diamonds, aes(depth)) +
geom_density(stat = "density", na.rm = TRUE, size = 2) +
geom_freqpoly(stat = "density", color = "purple", size = 1.2, na.rm = TRUE)

ggplot(diamonds, aes(depth)) +
geom_density(stat = "bin", na.rm = TRUE, size = 2) +
geom_freqpoly(stat = "bin", color = "light blue", size = 1.2, na.rm = TRUE)

```

